---
title: "Burnout-Forschung"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# global RMarkdown options
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# load packages
library(MASS)
library(tidyverse)
library(corrplot)

# set global theme options for ggplot2
theme_update(panel.background = element_rect(fill = NA, color = "#153268"))
theme_update(legend.background = element_blank())
theme_update(panel.grid = element_blank())
theme_update(strip.background = element_rect(fill = "#153268"))
theme_update(strip.text = element_text(color = "white"))
theme_update(plot.background = element_rect(fill = "#f6f4f0"))
```

```{r}
simdata <- function(n = 500, r = 0.55, mu = c(0, 0), names = c("exh", "cyn")) {
  s = matrix(c(1, r, r, 1), ncol = 2)
  x <- matrix(rep(NA, times = n*2), ncol = 2)
  id <- rep(NA, times = n)
  
 if (is.vector(mu)) {
    x <- mvrnorm(mu = mu, Sigma = s, n = n)
    id <- 1:n
 } else if (nrow(mu) == n) {
   
    for (i in 1:n) {
      tmu <- c(mu[[i,1]], mu[[i, 2]])
      x[i,] <- mvrnorm(mu = tmu, Sigma = s, n = 1)
      id[i] <- i
    }  
  }
  colnames(x) <- names
  d <- as_tibble(x)
  d$id <- id
  
  return(d)
}
```

```{r}
sim_covariate <- function(n = 500, r = 0.6) {
  s = matrix(c(1, r, r^2, r, 1, r^2, r^2, r, 1), ncol = 3)
  x <- mvrnorm(mu = c(0, 0, 0), Sigma = s, n = n)
  colnames(x) <- c("t0", "t1", "t2")
  d <- as_tibble(x) %>% 
    mutate(id = 1:n()) %>% 
    pivot_longer(t0:t2, names_to = "time") %>% 
    arrange(time)
  return(d %>% pull("value"))
}
```

```{r}
set.seed(12345)
true <- simdata()
t1 <- simdata(mu = true)
t2 <- simdata(mu = true)
true$time <- "t0"
t1$time <- "t1"
t2$time <- "t2"
```


```{r}
corrdat <- bind_rows(true, t1, t2) %>% 
  pivot_wider(names_from = "time", values_from = c("exh", "cyn")) %>% 
  dplyr::select(-id) 

corrmat <- corrdat %>% cor()

corrmat %>% round(2)  %>% knitr::kable()
```

```{r}
corrdat %>% psych::pairs.panels(hist.col = "skyblue", stars = TRUE)
```


```{r}
values <- bind_rows(true, t1, t2) %>% 
  
  group_by(time) %>% 
  mutate(median.exh = median(exh)) %>% 
  mutate(median.cyn = median(cyn)) %>% 
  ungroup() %>% 
  
  mutate(exh.status = ifelse(exh < median.exh, "low", "high")) %>% 
  mutate(cyn.status = ifelse(cyn < median.cyn, "low", "high")) %>% 
  mutate(exh.status = factor(exh.status, levels = c("low", "high"), 
                             ordered = TRUE)) %>%
  mutate(cyn.status = factor(cyn.status, levels = c("low", "high"),
                             ordered = TRUE)) %>% 
  
  mutate(pattern = case_when(
    (exh < median.exh) & (cyn < median.cyn) ~ "engagement",
    exh.status != cyn.status ~ "inconsistent",
    (exh > median.exh) & (cyn > median.cyn) ~ "burnout"
  ))

values %>% sample_n(10)  %>% knitr::kable()
```

```{r}
df <- values %>% 
  mutate(workload = sim_covariate()) %>% 
  mutate(control = sim_covariate()) %>% 
  mutate(reward = sim_covariate()) %>% 
  mutate(community = sim_covariate()) %>% 
  mutate(fairness = sim_covariate()) %>% 
  mutate(values = sim_covariate())
```

```{r}
jitter <- runif(n = nrow(df), min = -0.05, max = 0.05)

df <- df %>% 
  mutate(time.int = recode(time, "t0" = 1, "t1" = 2, "t2" = 3)) %>% 
  mutate(time.jit = time.int + jitter)
```

```{r}
status_wide <- df %>% 
  dplyr::select(id, time, exh.status, cyn.status) %>% 
  pivot_wider(names_from = "time", values_from = c("exh.status", "cyn.status"))

df <- left_join(df, status_wide, by = "id")

status_wide %>% sample_n(10)  %>% knitr::kable()
```
```{r}
pattern_wide <- df %>% 
  dplyr::select(id, time, pattern) %>% 
  pivot_wider(names_from = "time", values_from = pattern, names_glue="{.value}_{time}")

df <- left_join(df, pattern_wide, by = "id")

pattern_wide %>% sample_n(10)  %>% knitr::kable()
```


```{r}
plt.single <- function(df) {
  plt <- df %>% 
  ggplot() +
  aes(x = time.jit, y = y) +
  geom_line(aes(group = id, color = gr), alpha = 0.1) +
  geom_point(alpha = 0.1, aes(color = gr)) +
  # geom_hline(yintercept = medians$median.exh_t1, linetype = "dashed") +
  scale_x_continuous(limits = c(0.5, 3.5), breaks = 1:3, 
                     labels = c("True value", "t1", "t2")) +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  stat_summary(fun = "mean", 
               aes(x = time.int, group = gr), 
               geom = "line", size = 1) +
  stat_summary(fun = "mean", 
               aes(x = time.int, group = gr, shape = gr), 
               geom = "point", size = 3, fill = "white") +
  scale_shape_manual(values = c(22, 24)) +
  NULL
  
  return(plt)
}
```

```{r}
df %>% mutate(gr = exh.status_t1, y = exh) %>% 
  plt.single() +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Measurement", y = "Exhaustion", color = "Group at t1",
       title = "Development of Exhaustion", shape = "Group at t1")
```

```{r}
df %>% mutate(gr = cyn.status_t1, y = cyn) %>% 
  plt.single() +
  labs(x = "Measurement", y = "Cynicism", color = "Group at t1",
       title = "Development of Cynicsm")
```

```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = F) +
  facet_wrap(~time, labeller = label_both)
```

```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = F) +
  facet_wrap(~pattern_t0, labeller = label_both)
```
```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = F) +
  facet_wrap(~pattern_t1, labeller = label_both)
```

```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn, color = pattern_t1) +
  geom_point(alpha = 0.3) +
  facet_wrap(~time, labeller = label_both) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  labs(x = "Exhaustion", y = "Cynicsm", color = "Pattern at t1",
       title = "Development of pattern at t1")
```
```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn, color = pattern_t0) +
  geom_point(alpha = 0.3) +
  facet_wrap(~time, labeller = label_both) +
  scale_color_brewer(palette = "Set2") +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  labs(x = "Exhaustion", y = "Cynicsm", color = "Pattern at t1",
       title = "Development of pattern at t1")
```


```{r fig.height=8}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn, color = pattern_t1) +
  geom_point(alpha = 0.3) +
  facet_wrap(pattern_t1~time, labeller = label_both) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  labs(x = "Exhaustion", y = "Cynicsm", color = "Pattern at t1",
       title = "Development of pattern at t1")
```

```{r fig.height=8}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn, color = pattern_t0) +
  geom_point(alpha = 0.5) +
  facet_wrap(pattern_t1~time, labeller = label_both) +
  scale_color_brewer(palette = "Set2") +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  labs(x = "Exhaustion", y = "Cynicsm", color = "True pattern",
       title = "Development of patterns at t0 and t1")
```
```{r fig.height=8}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn, color = pattern_t0) +
  geom_point(alpha = 0.5) +
  facet_wrap(pattern_t0~time, labeller = label_both) +
  scale_color_brewer(palette = "Set2") +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  labs(x = "Exhaustion", y = "Cynicsm", color = "True pattern",
       title = "Development of pattern at t0")
```

```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn, color = pattern_t2) +
  geom_point(alpha = 0.4) +
  facet_wrap(pattern_t1~time, labeller = label_both) +
  scale_color_brewer(palette = "Dark2") +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  labs(x = "Exhaustion", y = "Cynicsm", color = "Pattern at t2",
       title = "Development of patterns at t1 and t2")
```

## Reasons for development of this pattern over time

- Measurement error
- Actual change in values of exhaustion and cynicsm (not the case in this simulation)

## Statistical analysis

- Target: People who are inconsistent at t1 and have a burnout or engagement status at t2
  
  - 6 comparisons in this subgroup

- Targets could have been many other questions

  - what is the factor that makes people change from burnout t1 to engagement t2 instead of swapping to inconsistent or staying at burnout? *6 comparisons*
  - what is the factor that makes people change from burnout to any other status? *6 comparisons*
  
  - what is the factor that makes people change from engagement t1 to burnout t2 instead of swapping to inconsistent or staying at engagement? *6 comparisons*
  - what is the factor that makes people change from engagement to any other status? *6 comparisons*
  
  - what is the factor that makes people stay inconsistent instead of changing to any other status? *6 comparisons*
  
  - That makes 36 comparisons at t2 in total. Even more, if both inconsistent states are counted individually.
  - By using ANOVA, the number can be reduced, but is still at 6 * 3 = 18
  - Even more tests, if the covariates are looked at both at t1 and t2 (the number doubles)

  - general prediction of engagement or cynicsm via covariates
  - incorporation of third burnout criterium


```{r}
df %>% 
  group_by(pattern, time) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = time, values_from = n, 
              names_glue = "{.value}_{time}")  %>% 
  knitr::kable()

```

```{r}
dev_pattern <- df %>% 
  dplyr::select(id, time, pattern, exh.status_t1, exh.status_t2, cyn.status_t1, cyn.status_t2) %>% 
  pivot_wider(names_from = "time", names_glue = "{.value}_{time}",
  values_from = "pattern") %>% 
  group_by(exh.status_t1, exh.status_t2, cyn.status_t1, cyn.status_t2, 
           pattern_t1, pattern_t2) %>% 
  summarise(n = n()) %>% ungroup()
  
dev_pattern  %>% knitr::kable()
```

```{r}
dev_pattern %>% group_by(pattern_t1, pattern_t2) %>% 
  summarise(n = sum(n))  %>% knitr::kable()
```

```{r}
dev_pattern %>% group_by(exh.status_t1, exh.status_t2) %>% 
  summarise(n = sum(n))  %>% knitr::kable()
```

```{r}
dev_pattern %>% group_by(cyn.status_t1, cyn.status_t2) %>% 
  summarise(n = sum(n))  %>% knitr::kable()
```



```{r}
dif.df <- df %>% 
  mutate(dif = exh - cyn) %>% 
  group_by(time) %>% 
  mutate(dif.med = median(dif)) %>% 
  mutate(dif.q25 = quantile(dif, 0.25)) %>% 
  mutate(dif.q75 = quantile(dif, 0.75)) %>% 
  ungroup() %>% 

  mutate(dif.status1 = ifelse(dif < dif.med, "low", "high")) %>% 
  mutate(dif.status = case_when(
    dif < dif.q25 ~ "low",
    (dif.q25 < dif) & (dif < dif.q75) ~ "medium",
    dif.q75 < dif ~ "high"
    ))

dif.status_t1 <- dif.df %>% 
  filter(time == "t1") %>% 
  select(id, dif.status) %>% 
  rename(dif.status_t1 = dif.status)
  
dif.df <- left_join(dif.df, dif.status_t1, by = "id")

dif.t1 <- dif.df %>% 
  filter(time == "t1") %>% 
  select(id, dif) %>% 
  rename(dif.t1 = dif)

dif.df <- left_join(dif.df, dif.t1, by = "id")

dif.df %>% dplyr::select(id, time, dif, pattern, dif.med, dif.status)
```

```{r}
pattern_dif <- dif.df %>% 
  group_by(pattern_t1, dif.status_t1) %>% 
  summarise(n = n()) %>% ungroup()
 
pattern_dif
```

```{r}
pattern_dif %>% 
  mutate(dif_extreme = ifelse(dif.status_t1 == "high" | dif.status_t1 == "low","yes", "no")) %>% 
  group_by(dif_extreme, pattern_t1) %>% 
  summarise(n = sum(n))
```


```{r}
plt.dif_combined <- dif.df %>% 
  ggplot() +
  aes(x = time.jit, y = dif) +
  geom_line(aes(group = id, color = dif.t1), alpha = 0.2) +
  geom_point(alpha = 0.1) +
  scale_color_gradient2(low = "#0571b0", high = "#ca0020", mid = "#ffffbf") +
  scale_x_continuous(breaks = 1:3, labels = c("True value", "t1", "t2")) +
  labs(x = "Measurement", y = "Difference: Exh. - Cyn.") +
  theme(legend.position = "bottom") +
  NULL
```

```{r}
plt.dif_combined + 
  stat_summary(fun = "mean", geom = "line",
               aes(x = time.int, group = dif.status_t1)
               ) +
  stat_summary(fun = "mean", geom = "point", 
               size = 3,
               fill = "white",
               aes(shape = dif.status_t1, x = time.int)) +
  # scale_shape_manual(values = c(22, 24))
  scale_shape_manual(values = c(21, 22, 24))
```

```{r}
plt.dif_combined + 
  facet_wrap(~dif.status_t1, labeller = label_both) +
  stat_summary(fun = "mean", geom = "line",
               aes(x = time.int, group = dif.status_t1)
               ) +
  stat_summary(fun = "mean", geom = "point", 
               size = 3,
               fill = "white",
               aes(shape = dif.status_t1, x = time.int)) +
  scale_shape_manual(values = c(21, 22, 24)) +
  # scale_shape_manual(values = c(21, 22, 24)) +
  NULL
```
