---
title: "Burnout-Forschung"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# global RMarkdown options
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# load packages
library(MASS)
library(tidyverse)
library(corrplot)
```

```{r}
simdata <- function(n = 500, r = 0.55, mu = c(0, 0), names = c("exh", "cyn")) {
  s = matrix(c(1, r, r, 1), ncol = 2)
  x <- matrix(rep(NA, times = n*2), ncol = 2)
  id <- rep(NA, times = n)
  
 if (is.vector(mu)) {
    x <- mvrnorm(mu = mu, Sigma = s, n = n)
    id <- 1:n
 } else if (nrow(mu) == n) {
   
    for (i in 1:n) {
      tmu <- c(mu[[i,1]], mu[[i, 2]])
      x[i,] <- mvrnorm(mu = tmu, Sigma = s, n = 1)
      id[i] <- i
    }  
  }
  colnames(x) <- names
  d <- as_tibble(x)
  d$id <- id
  
  return(d)
}
```

```{r}
sim_covariate <- function(n = 500, r = 0.6) {
  s = matrix(c(1, r, r^2, r, 1, r^2, r^2, r, 1), ncol = 3)
  x <- mvrnorm(mu = c(0, 0, 0), Sigma = s, n = n)
  colnames(x) <- c("t0", "t1", "t2")
  d <- as_tibble(x) %>% 
    mutate(id = 1:n()) %>% 
    pivot_longer(t0:t2, names_to = "time") %>% 
    arrange(time)
  return(d %>% pull("value"))
}
```

```{r}
true <- simdata()
t1 <- simdata(mu = true)
t2 <- simdata(mu = true)
true$time <- "t0"
t1$time <- "t1"
t2$time <- "t2"
```


```{r}
corrdat <- bind_rows(true, t1, t2) %>% 
  pivot_wider(names_from = "time", values_from = c("exh", "cyn")) %>% 
  dplyr::select(-id) 

corrmat <- corrdat %>% cor()

corrmat %>% round(2)
```

```{r}
corrdat %>% psych::pairs.panels(hist.col = "skyblue", stars = TRUE)
```


```{r}
values <- bind_rows(true, t1, t2) %>% 
  
  group_by(time) %>% 
  mutate(median.exh = median(exh)) %>% 
  mutate(median.cyn = median(cyn)) %>% 
  ungroup() %>% 
  
  mutate(exh.status = ifelse(exh < median.exh, "low", "high")) %>% 
  mutate(cyn.status = ifelse(cyn < median.cyn, "low", "high")) %>% 
  mutate(exh.status = factor(exh.status, levels = c("low", "high"), 
                             ordered = TRUE)) %>%
  mutate(cyn.status = factor(cyn.status, levels = c("low", "high"),
                             ordered = TRUE)) %>% 
  
  mutate(pattern = case_when(
    (exh < median.exh) & (cyn < median.cyn) ~ "engagement",
    exh.status != cyn.status ~ "inconsistent",
    (exh > median.exh) & (cyn > median.cyn) ~ "burnout"
  ))

values
```

```{r}
df <- values %>% 
  mutate(workload = sim_covariate()) %>% 
  mutate(control = sim_covariate()) %>% 
  mutate(reward = sim_covariate()) %>% 
  mutate(community = sim_covariate()) %>% 
  mutate(fairness = sim_covariate()) %>% 
  mutate(values = sim_covariate())
```

```{r}
jitter <- runif(n = nrow(df), min = -0.05, max = 0.05)

df <- df %>% 
  mutate(time.int = recode(time, "t0" = 1, "t1" = 2, "t2" = 3)) %>% 
  mutate(time.jit = time.int + jitter)

df
```

```{r}
status_wide <- df %>% 
  dplyr::select(id, time, exh.status, cyn.status) %>% 
  pivot_wider(names_from = "time", values_from = c("exh.status", "cyn.status"))

df <- left_join(df, status_wide, by = "id")

status_wide  
```
```{r}
pattern_wide <- df %>% 
  dplyr::select(id, time, pattern) %>% 
  pivot_wider(names_from = "time", values_from = pattern, names_glue="{.value}_{time}")

df <- left_join(df, pattern_wide, by = "id")

pattern_wide  
```


```{r}
df %>% 
  mutate(group.exh = exh.status_t1) %>% 
  ggplot() +
  aes(x = time.jit, y = exh) +
  geom_line(aes(group = id, color = group.exh), alpha = 0.1) +
  geom_point(alpha = 0.05) +
  scale_color_brewer(palette = "Set1") +
  # geom_hline(yintercept = medians$median.exh_t1, linetype = "dashed") +
  scale_x_continuous(limits = c(0.5, 3.5), breaks = 1:3, 
                     labels = c("True value", "t1", "t2")) +
  labs(x = "Measurement", y = "Exhaustion", color = "Group at t1",
       title = "Development of Exhaustion") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  stat_summary(fun = "mean", 
               aes(x = time.int, group = group.exh), 
               geom = "line", size = 1) +
  stat_summary(fun = "mean", 
               aes(x = time.int, group = group.exh), 
               geom = "point", size = 3) +
  NULL
```
```{r}
df %>% 
  mutate(group.cyn = cyn.status_t1) %>% 
  ggplot() +
  aes(x = time.jit, y = cyn) +
  geom_line(aes(group = id, color = group.cyn), alpha = 0.1) +
  geom_point(alpha = 0.05) +
  # scale_color_brewer(palette = "Set2") +
  # geom_hline(yintercept = medians$median.exh_t1, linetype = "dashed") +
  scale_x_continuous(limits = c(0.5, 3.5), breaks = 1:3, 
                     labels = c("True value", "t1", "t2")) +
  labs(x = "Measurement", y = "Cynicism", color = "Group at t1",
       title = "Development of Cynicsm") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  stat_summary(fun = "mean", 
               aes(x = time.int, group = group.cyn), 
               geom = "line", size = 1) +
  stat_summary(fun = "mean", 
               aes(x = time.int, group = group.cyn), 
               geom = "point", size = 3) +
  NULL
```

```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = F) +
  facet_wrap(~time, labeller = label_both)
```

```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = F) +
  facet_wrap(~pattern_t0, labeller = label_both)
```
```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = F) +
  facet_wrap(~pattern_t1, labeller = label_both)
```

```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn, color = pattern_t1) +
  geom_point(alpha = 0.3) +
  facet_wrap(~time, labeller = label_both) +
  facet_wrap(pattern_t1~time, labeller = label_both) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  labs(x = "Exhaustion", y = "Cynicsm", color = "Pattern at t1",
       title = "Development of pattern at t1")
```

```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn, color = pattern_t0) +
  geom_point(alpha = 0.5) +
  facet_wrap(pattern_t1~time, labeller = label_both) +
  scale_color_brewer(palette = "Set2") +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  labs(x = "Exhaustion", y = "Cynicsm", color = "True pattern",
       title = "Development of patterns at t0 and t1")
```
```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn, color = pattern_t0) +
  geom_point(alpha = 0.5) +
  facet_wrap(pattern_t0~time, labeller = label_both) +
  scale_color_brewer(palette = "Set2") +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  labs(x = "Exhaustion", y = "Cynicsm", color = "True pattern",
       title = "Development of pattern at t0")
```
```{r}
df %>% 
  ggplot() +
  aes(x = exh, y = cyn, color = pattern_t2) +
  geom_point(alpha = 0.4) +
  facet_wrap(pattern_t1~time, labeller = label_both) +
  scale_color_brewer(palette = "Dark2") +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(override.aes = list(alpha = 1, size=2))) +
  labs(x = "Exhaustion", y = "Cynicsm", color = "Pattern at t2",
       title = "Development of patterns at t1 and t2")
```

## Reasons for development of this pattern over time

- Measurement error
- Actual change in values of exhaustion and cynicsm (not the case in this simulation)

## Statistical analysis

- Target: People who are inconsistent at t1 and have a burnout or engagement status at t2
  
  - 6 comparisons in this subgroup

- Targets could have been many other questions

  - what is the factor that makes people change from burnout t1 to engagement t2 instead of swapping to inconsistent or staying at burnout? *6 comparisons*
  - what is the factor that makes people change from burnout to any other status? *6 comparisons*
  
  - what is the factor that makes people change from engagement t1 to burnout t2 instead of swapping to inconsistent or staying at engagement? *6 comparisons*
  - what is the factor that makes people change from engagement to any other status? *6 comparisons*
  
  - what is the factor that makes people stay inconsistent instead of changing to any other status? *6 comparisons*
  
  - That makes 36 comparisons at t2 in total. Even more, if both inconsistent states are counted individually.
  - By using ANOVA, the number can be reduced, but is still at 6 * 3 = 18
  - Even more tests, if the covariates are looked at both at t1 and t2 (the number doubles)

  - general prediction of engagement or cynicsm via covariates
  - incorporation of third burnout criterium


## Sum score




```{r eval=FALSE, include=FALSE}
df2 <- df %>% 
  dplyr::select(id, exh.status, cyn.status, time) %>% 
  filter(time != "t0") %>% 
  pivot_wider(names_from = "time", 
              names_sep = ".",
              values_from = c("exh.status", "cyn.status"))

df3 <- df2 %>% 
  mutate(pattern.t1 = ifelse(exh.status.t1 == cyn.status.t1, 
                             "consistent", "inconsistent")) %>% 
  mutate(pattern.t2 = ifelse(exh.status.t2 == cyn.status.t2, 
                             "consistent", "inconsistent")) %>% 
  mutate(ex.change = ifelse(exh.status.t1 == exh.status.t2, 
                          "unchanged", "changed")) %>% 
  mutate(cyn.change = ifelse(cyn.status.t1 == cyn.status.t2, 
                             "unchanged", "changed")) %>% 
  mutate(pattern.change = ifelse(pattern.t1 == pattern.t2,
                                 "unchanged", "changed")) %>% 
  mutate(burnout.t1 = ifelse(exh.status.t1 == "high" & cyn.status.t1 == "high",
                             TRUE, FALSE)) %>% 
  mutate(burnout.t2= ifelse(exh.status.t2 == "high" & cyn.status.t2 == "high",
                             TRUE, FALSE)) %>% 
    mutate(engagement.t1= ifelse(exh.status.t1 == "low" & cyn.status.t1 == "low",
                             TRUE, FALSE)) %>% 
  mutate(engagement.t2= ifelse(exh.status.t2 == "low" & cyn.status.t2 == "low",
                             TRUE, FALSE)) %>% 
  mutate(inconsistent.t1= ifelse(exh.status.t1 != cyn.status.t1,
                             TRUE, FALSE)) %>% 
  mutate(inconsistent.t2 = ifelse(exh.status.t2 != cyn.status.t2,
                             TRUE, FALSE)) %>% 
  mutate(status.dev = case_when(
    (burnout.t1 == burnout.t2) & (burnout.t1 == TRUE) ~ "stable.burnout",
    (engagement.t1 == engagement.t2) & (engagement.t1 == TRUE) ~ "stable.engagement",
    (inconsistent.t1 == inconsistent.t2) & (inconsistent.t1 == TRUE) & (ex.change == "unchanged") ~ "stable.inconsistent",
    (inconsistent.t1 == inconsistent.t2) & (inconsistent.t1 == TRUE) & (ex.change == "changed" | cyn.change == "changed") ~ "swap.inconsistent",
    (pattern.t1 == pattern.t2) & (pattern.t1 == "consistent") & burnout.t1 == TRUE & burnout.t2 == FALSE ~ "fullswap.to_engagement",
    (pattern.t1 == pattern.t2) & (pattern.t1 == "consistent") & burnout.t1 == FALSE & burnout.t2 == TRUE ~ "fullswap.to_burnout",
    ((inconsistent.t1 == TRUE) & (inconsistent.t2 == FALSE)) & burnout.t2 ~ "swap.to_burnout",
    ((inconsistent.t1 == TRUE) & (inconsistent.t2 == FALSE)) & engagement.t2 ~ "swap.to_engagement",
    (inconsistent.t1 == FALSE) & (inconsistent.t2 == TRUE) ~ "swap.to_inconsistent"
  ))

df4 <- df3 %>% 
  # dplyr::select("id":"pattern.change", "status.dev") %>% 
  group_by(pattern.change, exh.status.t1, exh.status.t2, cyn.status.t1, cyn.status.t2, pattern.t1, pattern.t2, status.dev) %>% 
  summarise(n = n())
  
df4

df4 %>% 
  filter(cyn.status.t1 == "high", exh.status.t1 == "high") %>% 
  filter(cyn.status.t2 == "high", exh.status.t2 == "high") %>%
  pull(n)

df3 %>% group_by(cyn.status.t1, exh.status.t1, cyn.status.t2, exh.status.t2) %>% 
  summarise(n = n())
  
  filter(cyn.status.t1 == "low", exh.status.t1 == "low") %>% pull(n) %>% sum()

df4 %>% 
  filter(cyn.status.t1 == "high", exh.status.t1 == "low") %>% pull(n) %>% sum()

df4 %>% 
  filter(cyn.status.t1 == "low", exh.status.t1 == "high") %>% pull(n) %>% sum()

df4 %>% 
  filter(cyn.status.t1 == "high", exh.status.t1 == "high") %>% pull(n) %>% sum()

df4 %>% 
  filter(cyn.status.t2 == "low", exh.status.t2 == "low") %>% pull(n) %>% sum()

df4 %>% 
  filter(cyn.status.t2 == "high", exh.status.t2 == "low") %>% pull(n) %>% sum()

df4 %>% 
  filter(cyn.status.t2 == "low", exh.status.t2 == "high") %>% pull(n) %>% sum()

df4 %>% 
  filter(cyn.status.t2 == "high", exh.status.t2 == "high") %>% pull(n) %>% sum()
```






